Fast and Error-Adaptive Influence Maximization based on Count-Distinct Sketches

### 2022-10-07 paper.tex
progress

\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  \graphicspath{{./images/}}
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png,.svg}
\else 
\fi
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage[]{booktabs}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:~}}
\renewcommand{\algorithmicensure}{\textbf{Output:~}}
\newcommand\ovr[1]{\overrightarrow{#1}}
\newcommand\acro{{\sc{HyperFuseR\xspace}\xspace}\xspace}
\newcommand\kktodo[1]{\textcolor{red}{#1}}
\newcommand\ggx[1]{\textcolor{blue}{#1}}
\newcommand\fixme[1]{\textcolor{red}{#1}}
\newcommand\acrotwo{
\multirow{2}{*}{ \shortstack{ \sc{Hyp.}\\\sc{Fus.}} }
}
\setlength{\intextsep}{0pt}%

\usepackage{array}
\usepackage{color}
\usepackage{listings}
\usepackage{placeins}
\usepackage{enumitem}
\newcommand{\resultsize}{0.8}
\algnewcommand{\algorithmicforeach}{\textbf{for each}}
\algdef{SE}[FOR]{ForEach}{EndForEach}[1]
  {\algorithmicforeach\ #1\ \algorithmicdo}% \ForEach{#1}
  {\algorithmicend\ \algorithmicforeach}% \EndForEach

@@

@@

\begin{wrapfigure}{l}{0.45\linewidth}
    \begin{center}
      \includegraphics[width=1\linewidth]{images/fminit.pdf}
    \end{center}
    \caption{Toy sketch initialization of vertex $x$ with four registers and 4-bit hash functions.}
  \end{wrapfigure}
The method first performs a diffusion process; the process starts with per-vertex sketches that are initialized with the hash value of the corresponding vertex~(i.e., every vertex reaches to itself). Then, for all the (sampled) edges ($u, v$), i.e., for the ones that contribute to the diffusion for this simulation, the sketch of the source vertex $u$ is merged into that of the target vertex $v$ until all sketch registers for all the vertices converge. The merge operation utilized in this process is slightly different from the conventional one and retrofitted to mimic the IC diffusion.

\begin{wrapfigure}{l}{0.35\linewidth}
    \begin{center}
      \includegraphics[width=1\linewidth]{images/fmmerge.pdf}
    \end{center}
    \caption{Example of merging vertex $x$ and $y$ with four registers.}
  \end{wrapfigure}
Throughout the process, each register is used only for a single sample/simulation. 
For an edge ($u, v$) in simulation $j$ where $v$ is a live vertex, an update, i.e., a merge operation, on $u$'s register is performed on the corresponding register $M_u$[$j$], i.e., $M_u[j] = max(M_u[j],M_v[j])$.
%if and only if the edge exists in the sample and \kktodo{bunu da degistirdim}\ggx{biz bunu yapmiyoruz, ben burada kayboldum.} $u$ is {\em live} in the current iteration. 
%A merge operation between sketch $M_u$,$M_v$ of vertices $u$,$v$,is only performed on $j$th register if there is a live edge between $u,v$ in sample $j$.
That is, at each iteration, vertices (outgoing) neighbors' reachability sets are added to their sketches.
This recursive formulation of the influence iteratively relays the reachability information among the vertices, allowing us to estimate the marginal influence for all vertices very fast.
%At each iteration, the sketches of the existing reachable vertices are modified. \kktodo{onceki cumleyi de degistirdim - dogru mu} \ggx{her vertex başkasına reachable olabilir, son cumleyi çıkartabilirz}
%Instead of simulating the diffusion sample-by-sample, vertex-by-vertex, or path-by-path, all the influence that passes through the edges in a single step is processed together.

@@

 Algorithm~\ref{algo:main} first initializes the reachability sets of all vertices by adding the vertices themselves. 
 That is for all vertices $u$, its $j$th register is set to $M_u[j]=clz(h_j(u))$ meaning $R_{{G}_j}(u) = \{u\}$ where $G_j$ is the $j$th sampled graph. 
 Then, we perform the diffusion process on the sketch registers whose pseudocode is given in Algorithm~\ref{algo:diffusion-step}. The diffusion starts by adding all the vertices to the {\em live vertex set} $L$. 
 Then at each step, the incoming edges of the live vertices are processed. 
 For a vertex $u$, its sketch, $M_u$, is updated by merging the sketches $M_v$ of all live outgoing neighbors vertices $v \in {L} \cap \Gamma^+_{G}(u)$. For each such vertex $v$ and simulation $j$, the operation $M_u[j] = max(M_u[j], M_v[j])$ is performed. This approach can be seen as a bottom-up, i.e., reversed,  diffusion process where at each iteration, the cardinality information is pulled from vertices neighbors.
%  If a vertex $u$ is live, its outgoing edges $(u, v)$ s.t. $v \in \Gamma_G(v)$ are processed and the sketches of $v$ are merged into those of $u$. \kktodo{burada edgeleri tersten process ettigimizi aciklamamiz gerek} burasi yanlis
If any of $u$'s sketch registers changes during this operation, it is added to the live vertex set $L'$ of the next iteration. Once the incoming edges of all live vertices are processed, the iteration ends. 
% Figure~\ref{fig:hf-processing} shows how \acro performs two simulations at the same time using sketch registers.

@@

The traditional Greedy algorithm~\cite{kempe2003maximizing} processes the simulations one-by-one and computes the vertices' reachability sets. On the other hand, \acro efficiently performs multiple simulations at once in a single-step iteration (which processes distance-1 neighborhoods). Since each iteration relays one level of cardinality information, this step requires at most $d$ iterations where $d$ is the diameter of $G$. Note that when processed individually (as the Greedy algorithm), the $j$th simulation over the sampled subgraph $G_j$ would require only at most $d_j$ iterations, which is the diameter of $G_j$, and probably much smaller than $d$ considering the sampled subgraphs being highly sparse. However, $d$ is also a loose upper bound for \acro. A better one is {\tt max}$\{d_j: 1 \leq j \leq {\cal J}\} + 1$ where $d_j$ is the diameter of $G_j$. We have observed that during these simulations later iterations have only a minor impact on the marginal influence scores. Hence, to further reduce the overhead of concurrent simulation processing and avoid bottleneck simulations due to remaining perimeter structures (e.g., paths on the perimeter), we employ an early-exit threshold $\epsilon_c$ over the remaining live vertices ratio, which is expected to be very small when only one or two (bottleneck) simulations remain. That is if $|L'| \leq |V| \times \epsilon_c$ the diffusion process in Algorithm~\ref{algo:diffusion-step} stops. Otherwise, $L$ is set to $L'$, $L'$ is cleared, and the next iteration starts. We used $\epsilon_c = 0.02$ to make \acro faster while keeping its quality almost the same.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%NEW END
\section{Experimental Results}\label{sec:evaluation}
We performed the experiments on a server with an 16-core {\tt Intel Xeon E5-2620 v4} CPU, running at 2.10GHz, and 192GB memory. The Operating System on the server is {\tt Ubuntu 20.04.2 LTS} with  5.4.0-65 kernel. The algorithms are implemented using {\tt C++20}, and compiled with {\tt GCC 9.3.0} with {\tt "-Ofast"} and {\tt "-march=native"} optimization flags. Multi-thread parallelization was achieved with {\tt OpenMP} pragmas. {\tt AVX2} instructions are utilized by handcrafted code with vector intrinsics.

@@

\hline
\multirow{6}{*}{\scriptsize\rotatebox[origin=c]{90}{Directed}}&{\tt Epinions} & 75,880 & 508,838  & 6.71 & 14\\
&{\tt Friendster} & 65,608,366 & 1,806,067,135  & 27.53 & 32\\
&{\tt LiveJournal} & 4,847,571 & 68,993,773  & 14.23 & 16\\
&{\tt Pokec} & 1,632,803 & 30,622,564  & 18.75 & 11\\
&{\tt Slashdot0811} & 77,360 & 905,468  & 11.70 & 10\\
&{\tt Slashdot0902} & 82,168 & 948,464  & 11.54 & 11\\
% &{\tt Twitter} & 81,306 & 2,420,766 & 1.37 & 29.77 
\end{tabular}
}
\end{table}
\FloatBarrier
\subsection{Experiment Settings} \label{sec:expset}

@@

In this work, we propose a sketch-based Influence Maximization algorithm that employs fused sampling and error-adaptive rebuilding. We provide a fast implementation of the algorithm that utilizes multi-threading to exploit multiple cores. Also, we present a performance comparison with state-of-the-art IM algorithms on real-world datasets and show that \acro{} can be an order of magnitude faster while obtaining the same influence. 
%With the proposed method, \acro can achieve $\maxspeedupSKIM\times$ speedup against a sketch-based algorithm {\sc Skim} and $\maxspeedupIMM\times$ speedup against an approximation-based algorithm {\sc Imm} while producing high-quality seed sets. 
%\ggx{Bu kısım değişmeli}
In the future, we will extend our work to a distributed GPGPU setting to process graphs with billions of vertices and edges under a minute. 
%The main issue with \acro on such an architecture is the broadcast of sketch registers. When the vertices are partitioned into different computation units, resulting sketch updates should be communicated to all partitions. To overcome the issue and reduce communication bottlenecks, graph preprocessing and speculative diffusion strategies will be employed.

>>>>

\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  \graphicspath{{./images/}}
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png,.svg}
\else 
\fi
\usepackage{amsmath}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage[]{booktabs}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:~}}
\renewcommand{\algorithmicensure}{\textbf{Output:~}}
\newcommand\ovr[1]{\overrightarrow{#1}}
\newcommand\acro{{\sc{HyperFuseR\xspace}\xspace}\xspace}
\newcommand\kktodo[1]{\textcolor{red}{#1}}
\newcommand\ggx[1]{\textcolor{blue}{#1}}
\newcommand\fixme[1]{#1}
% \newcommand\fixme[1]{\textcolor{red}{#1}}
\newcommand\acrotwo{
\multirow{2}{*}{ \shortstack{ \sc{Hyp.}\\\sc{Fus.}} }
}
% \setlength{\intextsep}{1pt}%

\usepackage{array}
\usepackage{color}
\usepackage{listings}
\usepackage{placeins}
\usepackage{enumitem}
\newcommand{\resultsize}{0.6}
\algnewcommand{\algorithmicforeach}{\textbf{for each}}
\algdef{SE}[FOR]{ForEach}{EndForEach}[1]
  {\algorithmicforeach\ #1\ \algorithmicdo}% \ForEach{#1}
  {\algorithmicend\ \algorithmicforeach}% \EndForEach

@@

HyperFuser uses \emph{directed} fused sampling, utilizes \emph{sketches} for both influence estimation and seed selection, and a novel error-adaptive sketch rebuilding strategy. On the other hand, {\sc InFuseR}~\cite{infuser} employs {\emph {undirected}} hash-based fused sampling and CELF optimizations for seed selection.

@@

% \begin{wrapfigure}{l}{0.45\linewidth}
%     \begin{center}
%       \includegraphics[width=1\linewidth]{images/fminit.pdf}
%     \end{center}
%     \caption{Toy sketch initialization of vertex $x$ with four registers and 4-bit hash functions.}
%   \end{wrapfigure}

% The method first performs a diffusion process; the process starts with per-vertex sketches that are initialized with the hash value of the corresponding vertex~(i.e., every vertex reaches to itself). Then, for all the (sampled) edges ($u, v$), i.e., for the ones that contribute to the diffusion for this simulation, the sketch of the source vertex $u$ is merged into that of the target vertex $v$ until all sketch registers for all the vertices converge. The merge operation utilized in this process is slightly different from the conventional one and retrofitted to mimic the IC diffusion.

% \begin{wrapfigure}{l}{0.35\linewidth}
%     \begin{center}
%       \includegraphics[width=1\linewidth]{images/fmmerge.pdf}
%     \end{center}
%     \caption{\label{fmmerge} Example of union operation on sketches of vertices $x$ and $y$ with four registers.}
%   \end{wrapfigure}

Throughout the process, each register is used only for a single sample/simulation. 
For an edge ($u, v$) in simulation $j$ where $v$ is a live vertex, an update, i.e., a merge operation, on $u$'s register is performed on the corresponding register $M_u$[$j$], i.e., $M_u[j] = max(M_u[j],M_v[j])$. 
The toy example in Fig.~\ref{fig:fmops} shows how the sketch operations are done on a toy problem. In Fig.~\ref{fig:fmops}(a), the sketches, in a toy problem with 4 samples and 4-bit hash functions, are initialized; the id of vertex $u$ is hashed and leading zeros of the hash are written to registers. In Fig.~\ref{fig:fmops}(a), the merge operation is illustrated on registers of vertex $u$, $M_u$, and vertex $v$, $M_v$; merge operation takes element-wise maximum of registers to find union of the reachability sets. Fig.~\ref{fig:fmops} shows writing the merged reachability sets back to registers w.r.t sampling using a conditional move operation.

%if and only if the edge exists in the sample and \kktodo{bunu da degistirdim}\ggx{biz bunu yapmiyoruz, ben burada kayboldum.} $u$ is {\em live} in the current iteration. 
%A merge operation between sketch $M_u$,$M_v$ of vertices $u$,$v$,is only performed on $j$th register if there is a live edge between $u,v$ in sample $j$.

\begin{figure}[!ht]
    \begin{center}
    \includegraphics[width=\linewidth]{images/fmops.pdf}
    \caption{ (a) \label{fminit} Toy sketch initialization of vertex $x$ with four registers and 4-bit hash functions. (b) \label{fmmerge} Example of union operation on sketches of vertices $x$ and $y$ with four registers. (c) \label{fmwrite} Diffusion operation from vertex $v$ to vertex $u$ w.r.t sampling in $\hat{G}$. 
     }\label{fig:fmops} 
    \end{center}
\end{figure}

That is, at each iteration, vertices (outgoing) neighbors' reachability sets are added to their sketches.
This recursive formulation of the influence iteratively relays the reachability information among the vertices, allowing us to estimate the marginal influence for all vertices very fast.
%At each iteration, the sketches of the existing reachable vertices are modified. \kktodo{onceki cumleyi de degistirdim - dogru mu} \ggx{her vertex başkasına reachable olabilir, son cumleyi çıkartabilirz}
%Instead of simulating the diffusion sample-by-sample, vertex-by-vertex, or path-by-path, all the influence that passes through the edges in a single step is processed together.

% \begin{wrapfigure}{l}{0.35\linewidth}
%     \begin{center}
%       \includegraphics[width=1\linewidth]{images/fmwrite.pdf}
%     \end{center}
%     \caption{Example of writing merged sketch $M_{xy}$ in fig.\ref{fmmerge} back to $M_x$ w.r.t diffusion process.}
%   \end{wrapfigure}

@@

 Algorithm~\ref{algo:main} first initializes the reachability sets of all vertices by adding the vertices themselves. 
 That is for all vertices $u$, its $j$th register is set to $M_u[j]=clz(h_j(u))$ meaning $R_{{G}_j}(u) = \{u\}$ where $G_j$ is the $j$th sampled graph. 
 Then, we perform the diffusion process on the sketch registers whose pseudocode is given in Algorithm~\ref{algo:diffusion-step}. The diffusion starts by adding all the vertices to the {\em live vertex set} $L$. 
 Then at each step, the incoming edges of the live vertices are processed. 
 For a vertex $u$, its sketch, $M_u$, is updated by merging the sketches $M_v$ of all live outgoing neighbors vertices $v \in {L} \cap \Gamma^+_{G}(u)$. For each such vertex $v$ and simulation $j$, the operation $M_u[j] = max(M_u[j], M_v[j])$ is performed. This approach can be seen as a bottom-up, i.e., reversed,  diffusion process where at each iteration, the cardinality information is pulled from vertices neighbors.
 \begin{figure}[!ht]
    \begin{center}
    \includegraphics[width=\linewidth]{images/infuserdiff.pdf}
    \caption{InFuseR diffusion process illustrated for only vertex $1$ on toy sample graph.  (a) \label{fig:infuserinit} Toy graph InFuseR initialization.  (b) \label{infuserdiff1} First level of diffusion using \emph{push-based} method for vertex $1$. (c) Second level of diffusion following Fig.\ref{fig:infuserdiff}(a).
     }\label{fig:infuserdiff} 
    \end{center}
\end{figure}
 \begin{figure}[!ht]
    \begin{center}
    \includegraphics[width=\linewidth]{images/sketchdiff.pdf}
    \caption{\acro diffusion process illustrated for \emph{all} vertices on toy sample graph. (a)\label{hyperinit} Toy sketch initialization. (b) \label{hyperdiff1} First level of diffusion using \emph{Pull-based} method for \emph{all} vertices. (c) \label{hyperdiff2} Second level diffusion following Fig.\ref{fig:hyperdiff}(a), only updated vertices' sketches are pulled. 
     }\label{fig:hyperdiff} 
    \end{center}
\end{figure}

%  If a vertex $u$ is live, its outgoing edges $(u, v)$ s.t. $v \in \Gamma_G(v)$ are processed and the sketches of $v$ are merged into those of $u$. \kktodo{burada edgeleri tersten process ettigimizi aciklamamiz gerek} burasi yanlis
If any of $u$'s sketch registers changes during this operation, it is added to the live vertex set $L'$ of the next iteration. Once the incoming edges of all live vertices are processed, the iteration ends. 
% Figure~\ref{fig:hf-processing} shows how \acro performs two simulations at the same time using sketch registers.

@@

The traditional Greedy algorithm~\cite{kempe2003maximizing} processes the simulations one-by-one and computes the vertices' reachability sets. On the other hand, \acro efficiently performs multiple simulations at once in a single-step iteration (which processes distance-1 neighborhoods). Since each iteration relays one level of cardinality information, this step requires at most $d$ iterations where $d$ is the diameter of $G$. Note that when processed individually (as the Greedy algorithm), the $j$th simulation over the sampled subgraph $G_j$ would require only at most $d_j$ iterations, which is the diameter of $G_j$, and probably much smaller than $d$ considering the sampled subgraphs being highly sparse. However, $d$ is also a loose upper bound for \acro. A better one is {\tt max}$\{d_j: 1 \leq j \leq {\cal J}\} + 1$ where $d_j$ is the diameter of $G_j$. We have observed that during these simulations later iterations have only a minor impact on the marginal influence scores. Hence, to further reduce the overhead of concurrent simulation processing and avoid bottleneck simulations due to remaining perimeter structures (e.g., paths on the perimeter), we employ an early-exit threshold $\epsilon_c$ over the remaining live vertices ratio, which is expected to be very small when only one or two (bottleneck) simulations remain. That is if $|L'| \leq |V| \times \epsilon_c$ the diffusion process in Algorithm~\ref{algo:diffusion-step} stops. Otherwise, $L$ is set to $L'$, $L'$ is cleared, and the next iteration starts. We used $\epsilon_c = 0.02$ to make \acro faster while keeping its quality almost the same. Figure~\ref{fig:hyperdiff} shows a toy problem diffusion process on 4 vertices and a single sample, the process initializes sketches using respective hashes' leading zeros and bigger numbers are pulled from all recently changed neighbors. Whereas Infuser works vertices one by one, labeling visited vertices for the processed vertex. The process is repeated for multiple vertices until a seed vertex is found. Figure~\ref{fig:infuserdiff} shows an example in the same settings as Figure~\ref{fig:hyperdiff} to emphasize the difference between the two algorithms.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%NEW END
\section{Experimental Results}\label{sec:evaluation}
We performed the experiments on a server with an 16-core {\tt Intel Xeon E5-2620 v4} CPU, running at 2.10GHz, and 192GB memory. The Operating System on the server is {\tt Ubuntu 20.04.2 LTS} with  5.4.0-65 kernel. The algorithms are implemented using {\tt C++20}, and compiled with {\tt GCC 9.3.0} with {\tt "-Ofast"} and {\tt "-march=native"} optimization flags. Multi-thread parallelization was achieved with {\tt OpenMP} pragmas. {\tt AVX2} instructions are utilized by handcrafted code with vector intrinsics.
\\

@@

\hline
\multirow{6}{*}{\scriptsize\rotatebox[origin=c]{90}{Directed}}&{\tt Epinions} & 75,880 & 508,838  & 6.71 & 14\\
&{\tt Friendster} & 65,608,366 & 1,806,067,135  & 27.53 & 32\\
&{\tt LiveJournal} & 4,847,571 & 68,993,773  & 14.23 & 16\\
&{\tt Pokec} & 1,632,803 & 30,622,564  & 18.75 & 11\\
&{\tt Slashdot0811} & 77,360 & 905,468  & 11.70 & 10\\
&{\tt Slashdot0902} & 82,168 & 948,464  & 11.54 & 11\\
% &{\tt Twitter} & 81,306 & 2,420,766 & 1.37 & 29.77 
\end{tabular}
}
\end{table}
% \FloatBarrier
\subsection{Experiment Settings} \label{sec:expset}

@@

In this work, we propose a sketch-based Influence Maximization algorithm that employs fused sampling and a novel error-adaptive rebuilding strategy. We provide a fast implementation of the algorithm that utilizes multi-threading to exploit multiple cores. Also, we present a performance comparison with state-of-the-art IM algorithms on real-world datasets and show that \acro{} can be an order of magnitude faster while obtaining the same influence. 
%With the proposed method, \acro can achieve $\maxspeedupSKIM\times$ speedup against a sketch-based algorithm {\sc Skim} and $\maxspeedupIMM\times$ speedup against an approximation-based algorithm {\sc Imm} while producing high-quality seed sets. 
%\ggx{Bu kısım değişmeli}
In the future, we will extend our work to a distributed GPGPU setting to process graphs with billions of vertices and edges under a minute. 
%The main issue with \acro on such an architecture is the broadcast of sketch registers. When the vertices are partitioned into different computation units, resulting sketch updates should be communicated to all partitions. To overcome the issue and reduce communication bottlenecks, graph preprocessing and speculative diffusion strategies will be employed.
